Towards Semantic Patches

### 2010-06-18 papers/MIPS2010/abstract.tex
* Remove reference to metatheorems.
git-svn-id: http://gasp.dev.regis-gianas.com/svn/trunk@108 403e655e-980a-47bd-8c60-1ee5348bb6b7

The external layer offers views on the objects of the theory. A
view~$w$ is a human readable, and editable, partial representation of
an object~$t$. A view is obtained from an object through a
projection~$\pi$.  The user transforms the view $w$ into a view $w'$
which must be $\textit{put back}$ into the metatheory as a modified
object $t'$ using a $\textit{putback}$ function $\sigma(t, w)$.  In
concrete syntax, an object~$t$ may be denoted simply using its
syntactic definition but also, more interestingly, by a reference into
the repository or even as the application of a metatheorem.

@@

Computing the minimal edition script between two abstract syntax trees
is a NP-hard problem that has been heavily studied, in particular in
bio-informatics.  We guess that, in practice, we can restrict the
search to local transformations that preserve substructures (that
roughly consist in inserting, removing or moving around entire
sub-trees). In that case, an existing polynomial algorithm might be
used~\cite{REF!}. Besides, abstract trees are relatively small in
practice.

We are more concerned about our ability to elaborate high-level patches
from these low-level edition scripts. Actually, it is easy to deduce
from a low-level edition script, a low-level patch, based on object
syntax constructor elimination and introduction. However, such a
low-level patch may forbid a more efficient reuse of existing
objects. For instance, let us assume a transformation of an
object-level term~$x + x$ into~ he term~$2 + 2$. A low-level patch
would represents the transformation as the replacement of the first
sub-term $x$ by the sub-term~$2$ followed by the replacement of the
second sub-term $x$ by the sub-term~$2$. If we had an object-level type
derivation for $x + x$, the integration function would have to build
two sub-derivations, one for each replaced sub-terms.  Instead, a
high-level patch would represent that transformation as a substitution
of the free occurrences of~$x$ by the term $2$ in the term~$x + x$. In
that case, the integration function has the opportunity to apply a
metatheorem about the stability of typing derivations under
substitution and finally reuse the initial typing derivation
directly. Inferring automatically such a high-level patch seems an
unreachable objective~: there is clearly no most general high-level
patch. Nonetheless, an interactive process can be designed to help the
user building his high-level patch from a set of pre-computed
interpretations of the low-level patch.

@@

A more useful integration propagates changes to the context that uses
the modified object in order to trigger the updating this context
too. Again, in full generality, it is not possible to update the
context automatically. Instead, we choose to generate
\textit{challenges}, which are adaptation patches with holes. These
challenges can be addressed by the author that wrote the patch to be
integrated, by the author of the context or automatically when there
is no ambiguity (think of a context that uses a function whose
body has changed but not its type).

>>>>

The external layer offers views on the objects of the theory. A
view~$w$ is a human readable, and editable, partial representation of
an object~$t$. A view is obtained from an object through a
projection~$\pi$.  The user transforms the view $w$ into a view $w'$
which must be $\textit{put back}$ into the metatheory as a modified
object $t'$ using a $\textit{putback}$ function $\sigma(t, w)$.  In
concrete syntax, an object~$t$ may be denoted simply using its
syntactic definition but also, more interestingly, by a reference to this 
object into the repository.

@@

Computing the minimal edition script between two abstract syntax trees
is a NP-hard problem that has been heavily studied, in particular in
bio-informatics.  We guess that, in practice, we can restrict the
search to local transformations that preserve substructures (that
roughly consist in inserting, removing or moving around entire
sub-trees). In that case, an existing polynomial algorithm may be
used~\cite{Touzet05alinear}. Besides, abstract trees are relatively small in
practice.

We are more concerned about the fact that existing algorithms are not
aware of binders that occur in proof-terms. Indeed, some standard
edition scripts must be forbidden because they could break
well-scopedness. As a consequence, we will have to adapt standard
algorithms to only look for scope-preserving edition scripts.

%% We are more concerned about our ability to elaborate high-level patches
%% from these low-level edition scripts. Actually, it is easy to deduce
%% from a low-level edition script, a low-level patch, based on object
%% syntax constructor elimination and introduction. However, such a
%% low-level patch may forbid a more efficient reuse of existing
%% objects. For instance, let us assume a transformation of an
%% object-level term~$x + x$ into~ he term~$2 + 2$. A low-level patch
%% would represents the transformation as the replacement of the first
%% sub-term $x$ by the sub-term~$2$ followed by the replacement of the
%% second sub-term $x$ by the sub-term~$2$. If we had an object-level type
%% derivation for $x + x$, the integration function would have to build
%% two sub-derivations, one for each replaced sub-terms.  Instead, a
%% high-level patch would represent that transformation as a substitution
%% of the free occurrences of~$x$ by the term $2$ in the term~$x + x$. In
%% that case, the integration function has the opportunity to apply a
%% metatheorem about the stability of typing derivations under
%% substitution and finally reuse the initial typing derivation
%% directly. Inferring automatically such a high-level patch seems an
%% unreachable objective~: there is clearly no most general high-level
%% patch. Nonetheless, an interactive process can be designed to help the
%% user building his high-level patch from a set of pre-computed
%% interpretations of the low-level patch.

@@

A more useful integration propagates changes to the context that uses
the modified object in order to trigger the updating of this context
(which may trigger the updating of other tree, and so forth). Again,
in full generality, it is not possible to update the context
automatically. Instead, we choose to generate \textit{challenges},
which are adaptation patches with holes. These challenges can be
addressed by the author that wrote the patch to be integrated, by the
author of the context or automatically when there is no ambiguity
(think of a context that uses a function whose body has changed but
not its type).
